A script to take in the raw disturbance stack and create simplified versions of it:
1) Ignore 4 & 5 hot-drought thresholds
2) Fire only
3) Insect only

Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 10/25/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files
 
```{r, echo = FALSE, warning = FALSE, results = FALSE, message = FALSE}
# SETUP ----
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra", #New raster data package, documentation pdf here: https://cran.r-project.org/web/packages/terra/terra.pdf
                      "sf", #New vector data package
                      "mapview", #For quick interactive mapping
                      "here", #Relative path best practices
                      "landscapemetrics",  #Fragstats for R
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse! https://www.futureverse.org/packages-overview.html; https://henrikbengtsson.github.io/course-stanford-futureverse-2023/
                      "here", #relative path best practice
                      "tictoc") #time running of processes: tic("Name process")...toc()

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed


## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::here() #Check here location
here::i_am("workflow/stack-analysis/simple_disturbance_stack.Rmd")
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)
```

Set up directories and CRS

```{r}

## Set data directory
dataDir <- here::here("data")

## Set output directory & create if doesn't already exist --
outDir <- here::here("data", "derived")

# check if sub directory exists 
if (!dir.exists(outDir)){
  dir.create(here::here("data", "derived"))
}

#Set CRS
chosenCRS <- 'EPSG:5070'

#Set EPA L3 aoi (if set to NA, will operate on full stack)
aoiNm <- NA #"Southern Rockies"


```


Manage AOI

```{r}

if(!is.na(aoiNm)) {

#Load data for aoi
aoiL3 <- sf::st_read(here::here('data', 'aoi',  'NA_CEC_Eco_Level3.shp')) 

#Filter & merge to single vector
aoi <- aoiL3 %>%
  dplyr::filter(NA_L3NAME == aoiNm) %>%
  dplyr::group_by(NA_L3NAME) %>%
  dplyr::summarize(geometry = st_union(geometry)) 
}

```

Load disturbance data stack, clip and mask to AOI if relevant

```{r}

#Load disturbance stack & fix names
distStackFileNames <- list.files(here("data", "data_stack_forest_western_conus", "western-conus", "western-conus"),
                                   pattern = "*.tif$", full.names = TRUE)
distStack <- rast(distStackFileNames)
  
  
#Function to create disturbance layer names
re.name.stack <- function(dStack, dNamesRoot) {
  dNums <- as.character(seq(1999, 2020))
  dNames <- paste(dNamesRoot, dNums, sep="_")
  names(dStack) <- dNames
  return(dStack)
}

distStack <- distStack |> re.name.stack("forestDisturbance")


#Function to clip a raster to a vector, ensuring in same projection
#Returns raster in original projection, but clipped and masked to vector
careful.clip <- function(raster, vector) {
    if (sf::st_crs(vector) != terra::crs(raster)) { #if raster and vector aren't in same projection, change vector to match
      print("Projecting vector")
      vector <- sf::st_transform(vector, terra::crs(raster)) 
    } else {
      print("Vector already in raster CRS")
    }
    print("Clipping")
    r <- terra::crop(raster,
                     vector,
                     mask = TRUE) #crop & mask
    return(r)
}
  

#Clip & mask if AOI is set
if(!is.na(aoiNm)) {
  #Run function
  dStackAoi <- careful.clip(raster = distStack,
                                        vector = aoi)
  datatype(dStackAoi)
  writeNm <- paste0("dist_stack_", gsub(" ", "_", aoiNm), ".tif")
  terra::writeRaster(dStackAoi,
                     here::here(outDir, writeNm),
                     overwrite = TRUE,
                     datatype = 'INT1U')

} else if(is.na(aoiNm)) {
  dStackAoi <- distStack
}

if(is.na(aoiNm)) {
  aoiNm <- "west"
}


```

Original data stack codes are:
Value - landfire disturbance - hotter-drought
0 - none - no hotter-drought/fewer than 4 thresholds exceeded
1 - fire - no hotter-drought/fewer than 4 thresholds exceeded
2 - insect/disease - no hotter-drought/fewer than 4 thresholds exceeded
3 - other Landfire disturbance - no hotter-drought/fewer than 4 thresholds exceeded
4 - none - hotter-drought with 4 thresholds exceeded
5 - fire - hotter-drought with 4 thresholds exceeded
6 - insects/disease - hotter-drought with 4 thresholds exceeded
7 - other Landfire disturbance - hotter-drought with 4 thresholds exceeded
8 - none - hotter-drought with 5 thresholds exceeded
9 - fire - hotter-drought with 5 thresholds exceeded
10 - insects/disease - hotter-drought with 5 thresholds exceeded
11 - other Landfire disturbance - hotter-drought with 5 thresholds exceeded
12 - none - hotter-drought with 6 thresholds exceeded
13 - fire - hotter-drought with 6 thresholds exceeded
14 - insects/disease - hotter-drought with 6 thresholds exceeded
15 - other Landfire disturbance - hotter-drought with 6 thresholds exceeded



# Write out as combined raster

```{r}
terra::writeRaster(distStack,
                   here::here(outDir, "disturbance-stack-west.tif"),
                   overwrite = TRUE,
                   datatype = "INT1U",
                   gdal=c("COMPRESS=DEFLATE"))


```



Perform data stack simplification #1: Ignore 4 & 5-threshold hotter drought as well as "other" landfire disturbance

``` {r}

## Set new codes ----
oldCodes <- seq(0, 15)
newCodes <- c(0, 1, 2, 0, 0, 1, 2, 0, 0, 1, 2, 0, 3, 4, 5, 3)
# 0 - None
# 1 - Fire
# 2 - Insect/Disease
# 3 - Hot Drought
# 4 - Fire + Hot Drought
# 5 - Insect/Disease + Hot Drought

codes <- cbind(oldCodes, newCodes)


## Re-classify ----

#Classify to new codes
classifier <- matrix(codes, ncol=2, byrow=FALSE) #Two-column matrix can be used to classify integer values in a from -> to format


#Full stack
dStackAoiSimple <- dStackAoi %>% future.apply::future_lapply(FUN = function(x) {terra::classify(x, classifier)}) %>% rast()
dStackAoiSimple <- dStackAoiSimple |> re.name.stack("simple_disturbance")



#Write stack
writeNm <- paste0("simple_dist_stack_", gsub(" ", "_", aoiNm), ".tif")

terra::writeRaster(dStackAoiSimple,
                   here::here(outDir, writeNm),
                   overwrite = TRUE,
                   datatype = "INT1U")

```


Perform data stack simplification #2: Fire only

``` {r}

## Set new codes ----
oldCodes <- seq(0, 15)
newCodes <- c(0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0)
# 0 - None
# 1 - Fire


codes <- cbind(oldCodes, newCodes)


## Re-classify ----

#Classify to new codes
classifier <- matrix(codes, ncol=2, byrow=FALSE) #Two-column matrix can be used to classify integer values in a from -> to format

dStackAoiFireOnly <- dStackAoi %>% 
  future.apply::future_lapply(FUN = function(x) {terra::classify(x, classifier)}) %>% 
  terra::rast()
dStackAoiFireOnly <- dStackAoiFireOnly |> re.name.stack("fire_disturbance")

#Write stack
writeNm <- paste0("fire_dist_stack_", gsub(" ", "_", aoiNm), ".tif")

terra::writeRaster(dStackAoiFireOnly,
                   here::here(outDir, writeNm),
                   overwrite = TRUE,
                   datatype = "INT1U")


```

Perform data stack simplification #3: Insect only

``` {r}

## Set new codes ----
oldCodes <- seq(0, 15)
newCodes <- c(0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0)
# 0 - None
# 1 - Insect/Disease

codes <- cbind(oldCodes, newCodes)


## Re-classify ----

#Classify to new codes
classifier <- matrix(codes, ncol=2, byrow=FALSE) #Two-column matrix can be used to classify integer values in a from -> to format

dStackAoiInsectOnly <- dStackAoi %>% 
  future.apply::future_lapply(FUN = function(x) {terra::classify(x, classifier)}) %>% 
  terra::rast()
dStackAoiInsectOnly <- dStackAoiInsectOnly |> re.name.stack("insect_disturbance")

#Write stack
writeNm <- paste0("insect_dist_stack_", gsub(" ", "_", aoiNm), ".tif")

terra::writeRaster(dStackAoiInsectOnly,
                   here::here(outDir, writeNm),
                   overwrite = TRUE,
                   datatype = "INT1U")

```

Perform data stack simplification #4: Drought only

``` {r}

## Set new codes ----
oldCodes <- seq(0, 15)
newCodes <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1)
# 0 - None
# 1 - Drought

codes <- cbind(oldCodes, newCodes)


## Re-classify ----

#Classify to new codes
classifier <- matrix(codes, ncol=2, byrow=FALSE) #Two-column matrix can be used to classify integer values in a from -> to format

dStackAoiDroughtOnly <- dStackAoi %>% 
  future.apply::future_lapply(FUN = function(x) {terra::classify(x, classifier)}) %>% 
  terra::rast()
dStackAoiDroughtOnly <- dStackAoiDroughtOnly |> re.name.stack("drought_disturbance")

#Write stack
writeNm <- paste0("drought_dist_stack_", gsub(" ", "_", aoiNm), ".tif")

terra::writeRaster(dStackAoiDroughtOnly,
                   here::here(outDir, writeNm),
                   overwrite = TRUE,
                   datatype = "INT1U")

```

