Run basic spatial statistics on disturbance stack

Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 10/20/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

```{r, echo = FALSE, warning = FALSE}
# SETUP ----
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra", #New raster data package, documentation pdf here: https://cran.r-project.org/web/packages/terra/terra.pdf
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse! https://www.futureverse.org/packages-overview.html; https://henrikbengtsson.github.io/course-stanford-futureverse-2023/
                      "sf", #New vector data package
                      "mapview", #For quick interactive mapping
                      "here", #Relative path best practices
                      "landscapemetrics",  #Fragstats for R
                      "tictoc", #time running of processes
                      "tmap",
                      "glue", #easy strings
                      "remotes", #to access github libraries
                      "ggpmisc") #For adding R^2 to plots

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

#Install ecoregions package from github if needed, then load package
if(!"ecoregions" %in% rownames(installed.packages())) {
  remotes::install_github("tomroh/ecoregions")
}
library(ecoregions)


## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::here() #Check here location
here::i_am("code/SpatStats_01.Rmd")
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)
```

```{r, echo = FALSE}
#Load data
fileLocation <- here::here('data', 'nathan_outputs')

fireInsects <- terra::rast(here::here(fileLocation, "fire_and_insects.tif"))
fireStack <- terra::rast(here::here(fileLocation, "fire_stack.tif"))

#Load EPA ecoregion data from ecoregions package
epaL3 <- ecoregions::ContinentalUsEcoregion3 %>% 
  sf::st_transform(crs(fireInsects))
epaL4 <- ecoregions::ContinentalUsEcoregion4 %>% 
  sf::st_transform(crs(fireInsects))

  
#EPA ecoregion distinct name table for joins
nameJoin <- cbind(epaL4$us_l4name, 
                  gsub(" ", "", epaL4$us_l4name),
                  epaL4$us_l3name,
                  gsub(" ", "", epaL4$us_l3name)) %>% 
  `colnames<-`(c('us_l4name', 'us_l4nameclean', 'us_l3name', 'us_l3nameclean')) %>%
  as.data.frame() %>%
  dplyr::distinct() %>%
  dplyr::mutate(us_l4l3name = glue::glue("{us_l4nameclean}{us_l3nameclean}"))

```


Set the areas of interest

```{r}
aoiL3Interest <- epaL3 %>% filter(us_l3name == "Sierra Nevada" | 
                                us_l3name == "Middle Rockies" | 
                                us_l3name == "Idaho Batholith" | 
                                us_l3name == "Southern Rockies") %>%
  group_by(us_l3name) %>%
  summarize(geometry = st_union(geometry)) %>%
  mutate(us_l3nameclean = gsub(" ", "", us_l3name))


aoiL4Interest <- epaL4 %>%
  filter(us_l3name == "Sierra Nevada" | 
                                us_l3name == "Middle Rockies" | 
                                us_l3name == "Idaho Batholith" | 
                                us_l3name == "Southern Rockies") %>%
  group_by(us_l4name, us_l3name) %>%
  summarize(geometry = st_union(geometry)) %>%
  left_join(nameJoin, join_by(us_l3name, us_l4name))

```

```{r}
#Function to clip a raster to a vector, ensuring in same projection
#Returns raster in original projection, but clipped and masked to vector
careful.clip <- function(raster, vector) {
  if (sf::st_crs(vector) != terra::crs(raster)) { #if raster and vector aren't in same projection, change vector to match
    print("Projecting vector")
    vector <- sf::st_transform(vector, terra::crs(raster)) 
  } else {
    print("Vector already in raster CRS")
  }
  print("Clipping")
  r <- terra::crop(raster,
                   vector,
                   mask = TRUE) #crop & mask
  return(r)
}

#Function to clip a raster to a set of polygons (one clip per polygon in set),
#ensuring they are in the same projection
#Namefield indicates the field to use to name the item in the returned list (e.g. us_l3name)
#Returns the set of rasters as a named list 
careful.clip.set <- function(raster, vectors, namefield) {
  out <- list()
  nms <- c()
  for(i in 1:nrow(vectors)) {
    vec <- vectors[i,]
    clipped <- careful.clip(raster, vec)
    nm <- dplyr::select(sf::st_drop_geometry(vec) %>% ungroup(), {{namefield}}) %>%
  as.character()
    out[[nm]] <- clipped
  }
  return(out)
}

#fireInsectInterestList <- careful.clip.set(fireInsects, aoiL4Interest, us_l4l3name)
fireInsectInterestList <- careful.clip.set(fireInsects, aoiL3Interest, us_l3nameclean)



##################################
###GOT TO HERE IN SECOND RUN THROUGH CODE!
##################################



```

Now, for each raster in the list, visualize and export


```{r}

# numLand <- length(fireInsectInterestList)
# 
# if(numLand %% 2 == 0) {
#   nrow <- numLand %/% 2
# } else {
#   nrow <- (numLand %/% 2) + 1
# }
# ncol <- 2
# par(mfrow = c(nrow, ncol))
# 
# # Loop through the raster list and plot each raster
# for (name in names(fireInsectInterestList)) {
#   plot(fireInsectInterestList[[name]], main = name)
# }
# 
# # Reset the plotting layout
# par(mfrow = c(1, 1))
# 
# title('Fire+Insect')

outDir <- here("data", "intermediate", "l3_rast")
if (!dir.exists(outDir)){
  dir.create(outDir)
}

for(i in 1:length(fireInsectInterestList)) {
  filenm <- names(fireInsectInterestList)[[i]]
  filenm <- gsub(" ", "", filenm) %>%
    paste(. , ".tif", sep ="")
  land <- fireInsectInterestList[[i]]
  terra::writeRaster(land, here::here(outDir, filenm), datatype = "INT1U")
}

# nameJoin <- cbind(aoiL4Interest$US_L4NAME, 
#                   gsub(" ", "", aoiL4Interest$US_L4NAME)) %>% 
#   `colnames<-`(c('ogNm', 'flNm')) %>%
#   as.data.frame() %>%
#   left_join(y=aoiL4, join_by(ogNm == US_L4NAME)) %>%
#   select(ogNm, flNm, US_L3NAME) %>%
#   distinct()

rm(fireInsects, fireInsectInterestList)


```
```{r}
#add in forest mask layer and get forest area

fMask <- terra::rast(here::here('data', 'landfire-bps-derived-forest-mask.tif'))

#forestMaskL3 <- careful.clip.set(fMask, aoiL3Interest)
forestMaskL4 <- careful.clip.set(fMask, aoiL4Interest)

t <- terra::vect(sf::st_transform(aoiL4Interest, terra::crs(fMask)))

forestL4 <- forestMaskL4 %>% lapply(FUN = function(x) {global(x, fun = "notNA")})

forestL4Stats <- cbind(names(forestMaskL4),
                bind_rows(forestL4)) %>%
  mutate(forestAreaHa = notNA * 900 * .0001) %>% #900m^2 / pixel, ha conversion .0001
  left_join(nameJoin, join_by(`names(forestMaskL4)` == ogNm)) %>%
  filter(flNm != "AlpineZone") %>%
  filter(US_L3NAME == 'Middle Rockies' |
           US_L3NAME == 'Southern Rockies' |
           US_L3NAME == 'Idaho Batholith' |
           US_L3NAME == 'Sierra Nevada')

forestL3Stats <- forestL4Stats %>%
  group_by(US_L3NAME) %>%
  summarize(forestAreaHa = sum(forestAreaHa))

```

```{r}


# t <- fireInsectInterestList[[2]] #small
# t2 <- fireInsectInterestList[[9]] #small
# t3 <- fireInsectInterestList[[11]] #Southern forested, big landscape
# t4 <- fireInsectInterestList[[6]] #small area, many component polygons
# 
# tList <- list(fireInsectInterestList$`Eastern Batholith`, fireInsectInterestList$`Lochsa-Selway-Clearwater Canyons`)
# names(tList) <- c("Eastern Batholith", "Lochsa-Selway-Clearwater Canyons")

calc.class.metrics <- function(xLandscape) {
  classMetrics <- calculate_lsm(xLandscape,
              what = c(
                       'lsm_c_ed', #patch area-weighted edge density
                       'lsm_c_ai',
                       'lsm_c_clumpy',
                       'lsm_c_cpland',
                       'lsm_c_dcad',
                       'lsm_c_division',
                       'lsm_c_lpi',
                       'lsm_c_lsi',
                       'lsm_c_mesh',
                       'lsm_c_np',
                       'lsm_c_pafrac',
                       'lsm_c_tca',
                       'lsm_c_cohesion'), #aggregation of patches
              directions = 8,
              progress = TRUE)
  print("Class Metrics Done")
  gc()
  return(classMetrics)
}

calc.patch.metrics <- function(xLandscape) {
  patchMetrics <- calculate_lsm(xLandscape, what = "patch")
  print("Patch Metrics Done")
  gc()
  return(patchMetrics)
}

#metrics <- fireInsectInterestListClean %>% lapply(calc.metrics)

# tic()
# tt <- calc.class.metrics(t) #22 seconds
# toc()
# 
# tic()
# t2t <- calc.class.metrics(t2) #28
# toc()
# 
# 
# tic()
# ttPatch <- calc.patch.metrics(t) #86 seconds
# toc()
# 
# tic()
# ttc2 <- calc.class.metrics2(t) #160 seconds
# toc()
# 
# tic()
# tClass <- tList %>% lapply(calc.class.metrics) #50 seconds
# toc()
# 
# tic()
# tClass2 <- tList %>% lapply(calc.class.metrics) #66 sec
# toc()
# 
# tic()
# tPatch <- tList %>% lapply(calc.patch.metrics) #92 sec
# toc()
# 
# tic()
# t3t <- calc.class.metrics(t3) #161
# toc()
# 
# tic()
# t4t <- calc.class.metrics(t4) #145
# toc()



#check_landscape(fireInsectInterestList[[2]])

# idahoClass <- fireInsectInterestList %>% lapply(calc.class.metrics)
# idahoPatch <- fireInsectInterestList %>% lapply(calc.patch.metrics)

#l4Class <- fireInsectInterestList %>% lapply(calc.class.metrics)
#l4Patch <- fireInsectInterestList %>% lapply(calc.patch.metrics)

outDirT <- here("data", "intermediate", "l3_tbl")
if (!dir.exists(outDirT)){
  dir.create(outDirT)
}
if (!dir.exists(here::here(outDirT, "class_metrics"))){
  dir.create(here::here(outDirT, "class_metrics"))
}
if (!dir.exists(here::here(outDirT, "patch_metrics"))){
  dir.create(here::here(outDirT, "patch_metrics"))
}

inDir <- here("data", "intermediate", "l3_rast")

tifs <- list.files(inDir, pattern = ".tif", full.names = TRUE)
tifs <- tifs[3:4]

# outTbls <- list.files(here::here(outDir, "class_metrics"), pattern = ".csv") %>% tools::file_path_sans_ext()
# 
# indices_to_remove <- grep(paste(outTbls, collapse = "|"), tifs)
# 
# tifs <- tifs[-indices_to_remove]



calc.metrics.export <- function(tifs, outDir) {
    for(f in tifs) {
    land <- terra::rast(f)
    print(paste('Calculating metrics for', basename(f)))
    tic()
    cMetrics <- calc.class.metrics(land) %>%
      dplyr::left_join(y = lsm_abbreviations_names,
                       by = c("metric", "level")) %>%
      dplyr::mutate(flNm = f %>% 
        basename() %>% 
        gsub(".tif", "", .)) %>%
      dplyr::left_join(y = nameJoin,
                       by = c("flNm"))
    pMetrics <- calc.patch.metrics(land) %>%
      dplyr::left_join(y = lsm_abbreviations_names,
                       by = c("metric", "level")) %>%
      dplyr::mutate(flNm = f %>% 
        basename() %>% 
        gsub(".tif", "", .)) %>%
      dplyr::left_join(y = nameJoin,
                       by = c("flNm"))
    newNm <- f %>% 
      basename() %>% 
      gsub(".tif", ".csv", .)
    write_csv(cMetrics, here::here(outDir, "class_metrics", newNm))
    write_csv(pMetrics, here::here(outDir, "patch_metrics", newNm))
    rm(land, cMetrics, pMetrics)
    gc()
    toc()
  }
}

calc.class.metrics.export <- function(tifs, outDir) {
    for(f in tifs) {
    land <- terra::rast(f)
    print(paste('Calculating metrics for', basename(f)))
    tic()
    cMetrics <- calc.class.metrics(land) %>%
      dplyr::left_join(y = lsm_abbreviations_names,
                       by = c("metric", "level")) %>%
      dplyr::mutate(flNm = f %>% 
        basename() %>% 
        gsub(".tif", "", .)) %>%
      dplyr::left_join(y = nameJoin,
                       by = c("flNm"))
    newNm <- f %>% 
      basename() %>% 
      gsub(".tif", ".csv", .)
    write_csv(cMetrics, here::here(outDir, "class_metrics", newNm))
    rm(land, cMetrics)
    gc()
    toc()
  }
}

calc.class.metrics.export(tifs = tifs, outDir = outDirT)





```


plot the metrics:
-merge the outputs into one dataframe for each metric set
-bar plot of 

AREA

NUMBER OF PATCHES

PATCH AREA MEAN & MEDIAN

PATCH CORE AREA

ED: Area-weighted edge density. The metric describes the configuration of the
landscape, e.g. because an aggregation of the same class will result in a low edge density. The metric is standardized to the total landscape area, and therefore comparisons among landscapes with different total areas are possible. Increases from 0, ED=0 if only one patch

SHAPE: Shape metric. Equals SHAPE_MN = 0 if all patches are squares. Increases, without limit, as the shapes of patches become more complex. (better than perimiter-area ratios)

CONTIG: Shape metric. Equals the mean of the contiguity index on class level for all patches.  Asses the spatial connectedness (contiguity) of cells in patches. Ranges from 0 to 1

COHESION: COHESION is an ’Aggregation metric’. It characterises the connectedness of patches belonging to
class i. It can be used to asses if patches of the same class are located aggregated or rather isolated
and thereby COHESION gives information about the configuration of the landscape. Approaches COHESION = 0 if patches of class i become more isolated. Increases
if patches of class i become more aggregated.



```{r}

tblDir <- here::here('data', 'intermediate', 'l4_tbl')

cTbls <- list.files(here::here(tblDir, 'class_metrics'), pattern = ".csv", full.names = TRUE)
pTbls <- list.files(here::here(tblDir, 'patch_metrics'), pattern = ".csv", full.names = TRUE)

classStats <- lapply(cTbls, function(f){read.csv(f)}) %>% 
  bind_rows() %>%
  filter(class == 1)
woNmC <- classStats %>%
  filter(is.na(US_L3NAME)) %>%
  select(-US_L3NAME) %>%
  left_join(nameJoin)
classStats <- classStats %>%
  filter(!is.na(US_L3NAME)) %>%
  rbind(woNmC) %>%
  filter(US_L3NAME != "Northern Basin and Range") %>%
  mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))

patchStats <- lapply(pTbls, function(f){read.csv(f)}) %>% 
  bind_rows() %>%
  filter(class == 1)
woNmP <- patchStats %>%
  filter(is.na(US_L3NAME)) %>%
  select(-US_L3NAME) %>%
  left_join(nameJoin)
patchStats <- patchStats %>%
  filter(!is.na(US_L3NAME)) %>%
  rbind(woNmP) %>%
  filter(US_L3NAME != "Northern Basin and Range") %>%
  mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))


```

```{r}
#PATCH STATS

ggplot(patchStats, aes(x = flNm, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black")

ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 250))) +
  labs(title = "Patch Areas")

ggplot(patchStats %>% filter(name == "core area index"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "Core area index")

ggplot(patchStats %>% filter(name == "shape index"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "Shape index")

#CLASS STATS

ggplot(classStats) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")

ggplot(classStats %>% filter(name == "largest patch index")) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")

ggplot(classStats %>% filter(name == "largest patch index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 1))) +
  labs(title = "Largest patch index")

ggplot(classStats %>% filter(name == "core area percentage of landscape")) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")



#number of patches / area



ggplot(patchStats, aes(x = US_L3NAME, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "PatchStats All")

ggplot(classStats, aes(x = US_L3NAME, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "ClassStats All")


ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 25))) +
  labs(title = "Patch Areas") +
  theme(axis.text.x=element_blank())
ggsave(here::here('figs', 'patchAreasL3Lim.png'))


ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
#  coord_cartesian(xlim = NULL, ylim = (c(0, 2000))) +
  labs(title = "Patch Areas") +
  theme(axis.text.x=element_blank())
ggsave(here::here('figs', 'patchAreasL3NoLim.png'))

ggplot(patchStats %>% filter(name == "shape index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 20))) +
  labs(title = "Shape index")




ggplot(patchStats %>% filter(name == "core area index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 100))) +
  labs(title = "Core area index")
ggsave(here::here('figs', 'caIndexL3.png'))


#Function to filter a dataframe to only include values >= the selected percentile in each group
#Group is the column name to group by
#Value is the column with the values to calculate percentiles of
#Percentile is the percentile in the form e.g. 0.75
filter_by_percentile <- function(data, group, value, percentile) {
  result <- data %>%
    group_by({{group}}) %>%
    do(filter(., {{value}} >= quantile({{value}}, percentile))) %>%
    ungroup()
  return(result)
}


#Examine core area extremes

percentiles <- c(.8, .85, .9, .95, .975, .99)

patchAreaExtremes <- data.frame()

for(p in percentiles) {
  patchAreaFilt <- filter_by_percentile(data = patchStats %>% filter(name == "core area"),
                                    group = US_L3NAME,
                                    value = value,
                                    percentile = p) %>%
    mutate(Percent = p * 100)
  patchAreaExtremes <- rbind(patchAreaExtremes, patchAreaFilt)
}

ggplot(patchAreaExtremes, aes(x = US_L3NAME,
                          y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun = "mean", geom="point", shape=20, size=3, color="black", fill="black") +
  facet_wrap(~Percent) +
  labs(title = "Patch Core Area, top percentiles (no Y limit)") +
  xlab("EPA Lvl 3 Ecoregion") +
  theme(axis.text.x=element_blank())

ggplot(patchAreaExtremes, aes(x = US_L3NAME,
                          y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun = "mean", geom="point", shape=20, size=3, color="black", fill="black") +
  facet_wrap(~Percent) +
  coord_cartesian(xlim = NULL, ylim = (c(0, 1500))) +
  labs(title = "Patch Core Area, top percentiles (Y: 0-1500)") +
  xlab("EPA Lvl 3 Ecoregion") +
  theme(axis.text.x=element_blank())

ggsave(here::here('figs', 'patchAreaExtremes.png'))





# show_correlation(patchStats %>% filter(region == "Eastern Batholith") %>% select(-region), method = "pearson")
# 
# 
# show_correlation(patchStats %>% filter(region == "Lochsa-Selway-Clearwater Canyons") %>% select(-region), method = "pearson")


```

```{r}

classStatsReshape <- classStats %>%
  select(-type, -function_name, -metric) %>%
  pivot_wider(names_from = name, values_from = value)

patchStatsReshape <- patchStats %>%
  select(-type, -function_name, -metric) %>%
  pivot_wider(names_from = name, values_from = value)

patchSummary <- patchStatsReshape %>%
  group_by(flNm) %>%
  summarise(medCorePSize = median(`core area`),
            meanCorePSize = mean(`core area`),
            maxCorePSize = max(`core area`))



classStatsMap <- aoiL4Interest %>%
  left_join(nameJoin) %>%
  left_join(classStatsReshape,
            join_by(flNm == flNm,
                    US_L3NAME == US_L3NAME),
            keep = FALSE) %>%
  st_as_sf() %>%
  filter(flNm != "AlpineZone") %>%
  mutate(area = st_area(.)) %>%
  left_join(patchSummary) %>%
  left_join(forestL4Stats, by = join_by(flNm == flNm)) %>%
  mutate(nPatchPerFArea = `number of patches` / forestAreaHa,
        `core area percentage of forest` = (`total core area` / forestAreaHa) * 100)


tm <- tm_shape(classStatsMap) +
  tm_fill(col = "core area percentage of forest",
          colorNULL = "eeeeee",
          style = "cont") +
  tm_layout(title = "Core area percentage of the forest landscape")
tm
tmap_save(tm, here::here('figs', 'caPerL4Map.png'))

# tm <- tm_shape(classStatsMap %>% filter(!is.na(`core area percentage of forest`))) +
#   tm_fill(col = "core area percentage of forest",
#           colorNULL = "eeeeee",
#           style = "log10") +
#   tm_layout(title = "Core area percentage of the forest landscape")
# tm
# tmap_save(tm, here::here('figs', 'caPerL4MapLog10.png'))

tm <- tm_shape(classStatsMap %>% filter(`core area percentage of forest` < 2)) +
  tm_fill(col = "core area percentage of forest",
          colorNULL = "eeeeee",
          style = "cont") +
  tm_layout(title = "Core area percentage of the forest landscape <2%")
tm
tmap_save(tm, here::here('figs', 'caPerL4Map_noOutlier.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "number of patches", colorNULL = "eeeeee") +
  tm_layout(title = "number of patches")
tm
tmap_save(tm, here::here('figs', 'npatchL4Map.png'))


tm <- tm_shape(classStatsMap) +
  tm_fill(col = "nPatchPerFArea", colorNULL = "eeeeee") +
  tm_layout(title = "Number of patches per forest area")
tm
tmap_save(tm, here::here('figs', 'npatchpAreaL4Map.png'))

tm_shape(classStatsMap) +
  tm_fill(col = "largest patch index", colorNULL = "eeeeee") +
  tm_layout(title = "largest patch index")

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "medCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Median core size")
tm
tmap_save(tm, here::here('figs', 'medCorePL4Map.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "meanCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Mean core size")
tm
tmap_save(tm, here::here('figs', 'meanCorePL4Map.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "maxCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Max core size")
tm
tmap_save(tm, here::here('figs', 'maxCorePL4Map.png'))


```

```{r}

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = meanCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size")

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = medCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size")
ggsave(here::here('figs', 'MedianCoreSizeCAPF.png'))

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = maxCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Max core size")

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = nPatchPerFArea)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Number of patches per forest area")
  


patchStatsL4Top90Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .9) %>%
  group_by(flNm) %>%
  summarize(medCorePSize90 = median(`core area`),
            meanCorePSize90 = mean(`core area`))

patchStatsL4Top95Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .95) %>%
  group_by(flNm) %>%
  summarize(medCorePSize95 = median(`core area`),
            meanCorePSize95 = mean(`core area`))

patchStatsL4Top975Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .975) %>%
  group_by(flNm) %>%
  summarize(medCorePSize975 = median(`core area`),
            meanCorePSize975 = mean(`core area`))

classStatsMap2 <- classStatsMap %>% 
  left_join(patchStatsL4Top90Carea,
           join_by(flNm == flNm)) %>% 
    left_join(patchStatsL4Top95Carea,
           join_by(flNm == flNm)) %>%
  left_join(patchStatsL4Top975Carea,
           join_by(flNm == flNm))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize90)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 90 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF90.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize90)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 90 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize95)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 95 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF95.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize95)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 95 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize975)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 97.5 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF975.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize975)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 97.5 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = `largest patch index`)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Largest Patch Index")


```

```{r}
tm_shape(classStatsMap2) +
  tm_fill(col = "meanCorePSize90", colorNULL = "eeeeee") +
  tm_layout(title = "Mean core size, 90th percentile")

l3cap <- patchStatsReshape %>%
  group_by(US_L3NAME) %>%
  summarize(totCA = sum(`core area`)) %>%
  left_join(aoiL3Interest %>% 
    mutate(tAreaHa = drop_units(st_area(aoiL3Interest)) * .0001) %>%
    mutate(US_L3NAME = NA_L3NAME)) %>%
  mutate(`core area percentage of landscape` = (totCA / tAreaHa) * 100) %>%
  left_join(forestL3Stats) %>%
  mutate(`core area percentage of forest` = (totCA / forestAreaHa) * 100) %>%
    mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))

ggplot(l3cap) +
  geom_col(aes(x = US_L3NAME, y=`core area percentage of forest`, fill = US_L3NAME)) +
  labs(title = "Fire/Insect Core Area percentage of Forested (BPS) Landscape")
ggsave(here::here('figs', 'caPerL3.png'))

```